NURM: a tool to locate numerical differences in pipelines

### 2019-06-03 paper.tex
Edit figures and change the context

Figure~\ref{fig:overview} shows an overview of our method \tristan{add reprozip
tags to 1.b too. Add something to match ``(fixed)'' in 1.b -- cond 2}. For
each tested condition, the pipeline is containerized with Docker, and its
interface is described with Boutiques~\cite{glatard2017boutiques}.
\reprozip~\cite{Chirigati2016} is added to the container images for
provenance tracking. Our tool first clusters the subjects based on the
similarity of their process trees produced by \reprozip
(Fig.~\ref{fig:overview-cluster}) \tristan{``Hclustering'' isn't clear, just
mention ``clustering''}. This results in a reduced set of representative
process trees to evaluate. The processes in each representative process
tree are then labeled from the differences found in their outputs in the
two tested conditions (Fig.~\ref{fig:overview-classify}). To do that, a
difference file is first produced that contains all the differences between
conditions obtained by comparing file checksums
(Fig.~\ref{fig:overview-classify} - (A)). Using the difference file and the
process tree in Condition 1, we first identify temporary files and files
written by more than a single process -- these files result in undetermined
process labels \tristan{Add label in the Figure. Here and for any other step}.
We then label processes in 4 categories (Fig.~\ref{fig:overview-classify} -
(B)) depending whether they create differences (represented in red), remove
differences (blue), neither create nor remove differences (green), or are
undetermined (yellow). Process labelling is done in an iterative way,
through breakpoints and incremental updates in Condition 2, until no
undetermined process remains.

\begin{figure}
  \centering
  \begin{subfigure}{\columnwidth}
    \centering
    \includegraphics[width=.7\columnwidth]{images/Slide1}
    \caption{Subject clustering.}
    \label{fig:overview-cluster}
  \end{subfigure}
   \begin{subfigure}{\columnwidth}
    \centering
     \includegraphics[width=1\columnwidth]{images/Slide2}
     \caption{Process labelling.}
     \label{fig:overview-classify}
   \end{subfigure}
   \caption{Method overview}
   \label{fig:overview}
  \end{figure}

\subsection{Provenance capture and representation}
\tristan{All these sections should refer to labels on Figure 1.}

We use \reprozip to record: (1) the set of processes created by the
pipeline, using the \texttt{clone()} or \texttt{fork()} system call, and
(2) the set of files read, written and executed by each process, including
temporary files. \reprozip collects this information through the
\texttt{ptrace()} system call and stores it in a SQLite database.

@@

\begin{figure}
\centering
  \includegraphics[scale=0.25]{images/simple_graph}
  \caption{Process tree and graph
  constructed from the example pipeline in
  Listing~\ref{listing:sample-script}.
  Processes that read or write
  temporary files are 
  represented with squares. Plain edges 
  represent the process tree (\texttt{fork()} or \texttt{clone()} 
  system calls). Dashed edges represent file dependencies: temporary 
  files are in yellow and result files are in green.
  Every node in the tree is labeled using (1) a process id created by our
  reconstruction, (2) the name of the executable run by the process.
  Process 0 is the initial call to the \texttt{bash} interpreter, and
  processes 1, 3 and 4 are the calls to FSL bet, stats and maths made in
  Listing~\ref{listing:sample-script}. Process 2 is forked by process 1: it
  was captured by \reprozip while it did not appear in
  Listing~\ref{listing:sample-script}. \tristan{The script has only 1 call to fslstats, this should be fixed.}
}
  \label{fig:simple_script}
\end{figure}

@@

We label in four categories the processes of the representative graph of each subject:
\begin{enumerate}
\item Processes that read files that do not have differences and write files that do not have differences 
are labeled \emph{transparent} (Figure~\ref{fig:processes}.a).
\item Processes that read files 
that do not have differences but write files that have differences 
\emph{create} differences (Figure~\ref{fig:processes}.b).
\item Processes that read files 
that have differences and write files that do not have differences \emph{remove} 
differences (Figure~\ref{fig:processes}.c).
\item Processes that read files that have differences and write files that 
also have differences are called
\emph{undetermined} (Figure~\ref{fig:processes}.d).
\end{enumerate}

\begin{figure}%\centering
\centering
    \begin{subfigure}{0.4\linewidth}
        \includegraphics[scale=0.34]{images/green.png}
        \caption{Transparent}
        \label{fig:green}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.4\linewidth}
    \includegraphics[scale=0.34]{images/red.png}
    \caption{Creates differences}
    \label{fig:red}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.4\linewidth}
    \includegraphics[scale=0.34]{images/blue.png}
    \caption{Removes differences}
    \label{fig:blue}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.4\linewidth}
    \includegraphics[scale=0.34]{images/yellow.png}
    \caption{Undetermined}
    \label{fig:yellow}
\end{subfigure}
    \caption{Different type of classified process based on the input/output files.
  Dashed edges refer to the file dependencies between processes A and B 
  if a file written by A is read by B. Solid black edges refer to the 
  relationship between parent and child processes.}
    \label{fig:processes}
\end{figure}

Undetermined processes are due to temporary files, and files that were
written by more than a single process. In both situations, breakpoints must
be added to the pipeline to capture intermediary file states and label
undetermined processes.

we first capture temporary and overwritten files in condition 1
we set various 
breakpoints in the pipeline in condition 2 to check the output files 
produced by each process after running. We then label processes in 4 
categories. When processes that create differences are 
found, their written files in condition 2 are replaced (as explained 
later) and then re-label processes. This procedure (stop/resume) 
continues until the end of the pipeline execution when all processes 
are labeled correctly. \tristan{this looks completely out of place}

\tristan{Explain that condition 1 is re-executed after all these modifications are done?}

Another challenge is to label undetermined processes that read files with 
differences and write files with differences too. In such 
situations, it is not possible to determine from the pipeline results 
whether the process created differences, removed some differences, or 
was transparent.
To address this issue, we set breakpoints in the pipeline 
in Condition 2. Breakpoints are set by replacing all the pipeline 
processes with a custom script. This custom script labels each process 
as \emph{transparent}, \emph{create differences}, \emph{remove 
differences} or \emph{undetermined} immediately after running the original executable. 
The replacement is done through the PATH variable, as before. Moreover, 
this custom script copies the results obtained from the same process in 
Condition 1 to the pipeline output in Condition 2 if it is invoked with 
the arguments of a process that created differences. 
We developed an incremental approach that consists of the following steps:

\tristan{Not sure that we should call that breakpoints now.}

\tristan{Why do you need to modify condition 2 now, couldn't this be done on codition 1 too?}

\begin{enumerate}
  \item Run the pipeline in Condition 2; after running each process 
  label it as \emph{transparent}, \emph{create differences} or
  \emph{remove differences}.

  \item If there is \emph{create differences} processes then: 
  copies the results produced by the same process in condition 1 to the 
  pipeline output in Condition 2.

  \item Continue steps 1 and 2 until the end of the pipeline execution.
\end{enumerate}

This algorithm converges to a process tree without any 
\emph{undetermined} process at the end of the pipeline execution.

\tristan{Now that you're doing this, do you still need subject clustering? It looks like this isn't
much longer than regular execution.}

Figure~\ref{fig:iterations} illustrates our incremental labelling 
process for the example in Figure~\ref{fig:simple_script} \tristan{This is outdated, you should
update the figure and revise the text accordingly.}. At every 
step, processes that created differences are shown in red, processes that 
removed differences are in blue, processes that propagated differences are in 
yellow, and other processes (transparent processes) are in green. As in 
Figure~\ref{fig:simple_script}, plain black edges represent the process 
tree and dashed edges represent file dependencies: green edges 
represent files with no differences, while red edges represent files with 
differences. Temporary files are not represented because they have been 
backed up as previously described.

The three steps in Figure~\ref{fig:iterations} correspond to the 
steps of the labelling algorithm. At step one, \texttt{2\#bet2} 
is classified as difference creator (red) as it produced files with differences 
from files without differences. \texttt{3\#fslstats} is classified as difference  
remover (blue) as it produced files without differences from files with 
differences. \texttt{4\#fslmaths} and \texttt{5\#fslstats} are classified as 
unknown (yellow) as they produce files with differences from files with 
differences.

At step 2, the files produced by \texttt{2\#bet2} in the tested 
condition are replaced with the files produced by \texttt{2\#bet2} in 
the other condition. \texttt{3\#fslstats} is now classified as 
transparent, \texttt{4\#fslmaths} is now classified as difference creator 
and \texttt{5\#fslstats} is still unknown.

At step 3, the files produced by \texttt{4\#fslmaths} in the tested 
condition are replaced with the files produced by \texttt{4\#fslmaths} 
in the other condition. \texttt{5\#fslstats} is now transparent.

As a result of those 3 steps, the final process labelling is: 
\texttt{2\#bet2} and \texttt{4\#fslmath} are difference creators (red), 
\texttt{3\#fslstats} is difference remover (blue) and the other processes 
are transparent.

>>>>

Figure~\ref{fig:overview} shows an overview of our method. For
each tested condition, the pipeline is containerized with Docker, and its
interface is described with Boutiques~\cite{glatard2017boutiques}.
\reprozip~\cite{Chirigati2016} is added to one of the container images for
provenance tracking. 
Our tool starts by capturing temporary and multi write files in condition 1 
(Fig.~\ref{fig:overview-capturing}). 
To do that, the tool first executes pipeline on condition 1 to obtain process graph 
(Fig.~\ref{fig:overview-capturing} - (P)) and 
result files. Subsequently, it generates the list of processes that create these intermediary files. 
Using this list, the tool modifies docker image in condition 1 and exectutes the pipeline 
again through it, which turns to produce results including temporary and overwritten
files (Fig.~\ref{fig:overview-capturing} - (C1)).
In the second step, the processes are labeled from 
the differences found in their outputs in the
two tested conditions (Fig.~\ref{fig:overview-labelling}). To this order, 
the tool builds docker image for condition 2 by modifying pipeline processes that 
recognised from 
the process graph. Then, the modified docker image is executed in condition 2 and 
each process is labeled right 
after the running by comparing checksum of the results obtained in condition 1 
and output files of the process in condition 2 (Fig.~\ref{fig:overview-labelling} - (D)). 
We then label processes in 3 categories (Fig.~\ref{fig:overview-labelling} -
(L)) depending whether they create differences (represented in red) or not (green). 
Process labelling is done through hooks which are defined during the docker image modifying procedure, 
and incremental updates in condition 2, 
until the end of the pipeline execution.

\begin{figure}
  \centering
  \begin{subfigure}{\columnwidth}
    \centering
    \includegraphics[width=1\columnwidth]{images/Slide2}
    \caption{Capturing of temp and multi write files.}
    \label{fig:overview-capturing}
  \end{subfigure}
   \begin{subfigure}{\columnwidth}
    \centering
     \includegraphics[width=1\columnwidth]{images/Slide1}
     \caption{Process labelling.}
     \label{fig:overview-labelling}
   \end{subfigure}
   \caption{Method overview}
   \label{fig:overview}
  \end{figure}

\subsection{Provenance capture and representation}

We use \reprozip to create process graph (Fig.~\ref{fig:overview-capturing} - (P)), 
which records: (1) the set of processes created by the
pipeline, using the \texttt{clone()} or \texttt{fork()} system call, and
(2) the set of files read, written and executed by each process, including
temporary files. \reprozip collects this information through the
\texttt{ptrace()} system call and stores it in a SQLite database.

@@

\begin{figure}
\centering
  \includegraphics[scale=0.25]{images/simple_graph}
  \caption{Process tree and graph
  constructed from the example pipeline in
  Listing~\ref{listing:sample-script}.
  Processes that read or write
  temporary files are 
  represented with squares. Plain edges 
  represent the process tree (\texttt{fork()} or \texttt{clone()} 
  system calls). Dashed edges represent file dependencies: temporary 
  files are in yellow and result files are in green.
  Every node in the tree is labeled using (1) a process id created by our
  reconstruction, (2) the name of the executable run by the process.
  Process 0 is the initial call to the \texttt{bash} interpreter, and
  processes 1, 3 and 4 are the calls to FSL bet, stats and maths made in
  Listing~\ref{listing:sample-script}. Process 2 is forked by process 1: it
  was captured by \reprozip while it did not appear in
  Listing~\ref{listing:sample-script}. 
}
  \label{fig:simple_script}
\end{figure}

We label in two categories the processes of the graph of each subject:
\begin{enumerate}
\item Processes that read files that do not have differences and write files that do not have differences 
are labeled \emph{transparent} (Figure~\ref{fig:processes}.a).
\item Processes that read files 
that do not have differences but write files that have differences 
\emph{create} differences (Figure~\ref{fig:processes}.b).
\end{enumerate}

In addition, there are two more situations in labelling process as:
\begin{enumerate}
\item Processes that read files 
that have differences and write files that do not have differences \emph{remove} 
differences (Figure~\ref{fig:processes}.c).
\item Processes that read files that have differences and write files that 
also have differences are called
\emph{undetermined} (Figure~\ref{fig:processes}.d).
\end{enumerate}

\begin{figure}%\centering
\centering
    \begin{subfigure}{0.2\linewidth}
        \includegraphics[scale=0.3]{images/green.png}
        \caption{Transparent}
        \label{fig:green}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.2\linewidth}
    \includegraphics[scale=0.3]{images/red.png}
    \caption{Creates differences}
    \label{fig:red}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.2\linewidth}
    \includegraphics[scale=0.3]{images/blue.png}
    \caption{Removes differences}
    \label{fig:blue}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.2\linewidth}
    \includegraphics[scale=0.3]{images/yellow.png}
    \caption{Undetermined}
    \label{fig:yellow}
\end{subfigure}
    \caption{Different type of labeled process based on the input/output files.
  Dashed edges refer to the file dependencies between processes A and B 
  if a file written by A is read by B. Solid black edges refer to the 
  relationship between parent and child processes.}
    \label{fig:processes}
\end{figure}

Undetermined processes are due to temporary files, and files that were
written by more than a single process. In both situations, hooks must
be added to the pipeline to capture intermediary file states and label
undetermined processes.

After that all these modifications are done, we re-execute pipeline in condition 1.
Figure~\ref{fig:overview-capturing} - (C1) shows pipeline results including intermediary files 
captured as the outputs in condition 1.

Since all the result files are captured in condition 1, the tool 
is ready to label the piepeline processes based on the results in condition 2.
However, another challenge is to label undetermined processes that read files with 
differences and write files with differences too. In such 
situations, it is not possible to determine from the pipeline results 
whether the process created differences, removed some differences, or 
was transparent.
To address this issue, we set hooks in the pipeline 
in Condition 2. Hooks are set by replacing We identify dall the pipeline 
processes with a custom script. This custom script labels each process 
as \emph{transparent} or \emph{create differences} immediately after running 
the original executable. 
The replacement is done through the PATH variable, as before. Moreover, 
this custom script copies the results obtained from the same process in 
Condition 1 to the pipeline output in Condition 2 if 
it is invoked with 
the arguments of a process that created differences. 
Differences are identified by comparing process name and checksum of the output files.
We developed an incremental approach that consists of the following steps:

\begin{enumerate}
  \item Start the pipeline execution in Condition 2; 
        after hooking all the processes involved in the pipeline.
  \item Label each process as \emph{transparent} or \emph{create differences} 
        right after its running (Fig.~\ref{fig:overview-labelling} - (L)).
  \item Check if process is \emph{create differences} then: 
        copies the results produced by the same process in Condition 1 to the 
        pipeline output in Condition 2 (Fig.~\ref{fig:overview-labelling} - (M)). 
  \item Continue steps 2 and 3 until the end of the pipeline execution.
\end{enumerate}

This algorithm create a labeled process graph at the end of pipeline execution 
(Fig.~\ref{fig:overview-labelling} - (LG)).

Figure~\ref{fig:iterations} illustrates our incremental labelling 
process for the example in Figure~\ref{fig:simple_script}. At every 
step, processes that created differences are shown in red and other processes 
(transparent processes) are in green. As in 
Figure~\ref{fig:simple_script}, plain black edges represent the process 
tree and dashed edges represent file dependencies: green edges 
represent files with no differences, while red edges represent files with 
differences. Temporary files are not represented because they have been 
backed up as previously described.

The four steps in Figure~\ref{fig:iterations} correspond to the 
steps of the labelling algorithm. 
At step one, process \texttt{1\#bet} is executed and then labeled 
as transparent (green) as it produced files without differences.
At step 2, \texttt{2\#bet2} 
is labeled as difference creator (red) as it produced files with differences 
from files without differences. Therefore, the files produced by \texttt{2\#bet2} in  
condition 2 are replaced with the files produced by \texttt{2\#bet2} in 
condition 1.
At step 3, after running process \texttt{3\#fslstats}, it labeled as 
transparent as it write files without differences.
At step 4, process \texttt{4\#fslmaths} is labeled as difference creator 
as the last process of the pipeline.
As a result of these 4 steps, the final process labelling is: 
\texttt{2\#bet2} and \texttt{4\#fslmath} are difference creators (red) 
and the other processes are transparent.
It should be noted that process that remove differrences are not recognised 
by this method.

@@
